---
title: 'Statistical report: Use of the inhibitors of sodium-glucose cotransporter 2 in patients with heart failure and type 2 diabetes mellitus'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
header-includes:
   - \usepackage{draftwatermark}
---

\newpage 
\tableofcontents 
\listoffigures
\listoftables
\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SHFDB3, https://kiheartfailure.github.io/shfdb3/, v 3.1.2. 

## Inclusion/exclusion criteria

```{r flow}
mykable(flow, caption = "Flowchart")
```

First patient in: `r min(pdata$shf_indexdtm)` and last patient in: `r max(pdata$shf_indexdtm)`.  

QUESTION: Took LAST post (not first as stated in the SAP) 
since not index post anyway and last more probable to have sglt2. OK?

## Created variables 

The antidiabetic treatments were derived from the Dispensed Drug Registry where 
use was defined as dispensed drug within 5 months prior to index date (in SwedeHF) 
up until 14 days after index date. The following ATC codes were used to define the
respective treatments: 

```{r atc}
mykable(metalm, caption = "Treatments from DDR")
```

QUESTION: Should any of the A10BD be placed in another treat? 

QUESTION: ok with same time limit for other treats?

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Missing data

Missing data was imputed with multiple imputation using mice [@mice] for 10 
datasets and 10 iterations. Variables included in the model are indicated in 
Table \ref{tab:tab1}. The primary outcome, first heart failure hospitalization/CVD,
was included as the Nelson-Aalen estimator. Sglt2 type is not included in the imputation model.  

## Propensity scores

A propensity score for treatment with sglt2 was estimated for each patient with 
logistic regression for each of the 10 imputed datasets using with variables 
indicated in Table \ref{tab:tab1}. Continuous variables were modeled using 
cubic splines with 3 dfs. 1:2 matching without replacement [@match] 
was thereafter performed on the average of the resulting 10 ps [@psmatch]. Matching was 
allowed if the propensity score differed by 0.01 or less. The ability of the ps 
matching to balance the baseline characteristics was assessed by 
standardized mean differences (the difference in percentage between the means for 
the 2 groups divided by the mutual standard deviation). Standardized mean differences
of less than <0.1% are considered inconsequential. 

1:1 and 1:3 matching was also attempted but taking into consideration the no of 
patients retained 1:2 was deemed the best option
`r matchingn`.

QUESTION: Could abs try inverse prob weighing, but to be honest mi is a bloody problem so is nice
to be rid of it by using matching...

## Baseline characteristics

QUESTION: The covariates listed in the sap are not quite the same as presented in the table. Should I go by text or table?  
QUESTION: Just to check: You write you want Ischemic stroke: ICD-10: I63. This is different to general dataset were defined as I60-I64, I690-694; ICD-9: 430-434, 438A-E. Should be ONLY I63?

QUESTION: What is valvular interventions? Pci/cabg? So also included in ihd?

```{r, child = "../src/tab1.Rmd"}

```

## Associations with SGLT2 use

Odds ratios are estimated using logistic regression with SGLT2 use as outcome. 

Note to self: If keep cont vars present in units of 10 ect. 

Note to self: Add overall p for cat vars > 2 levels. 

QUESTION: Should make further selection of variables 
for adjusted model. This also applies to the ps model.
Models are not converging proparly...

```{r, child = "../src/ortab.Rmd"}

```

### Assumptions

Outliers were investigated with Cook's distance and multicollinearity 
with the variance inflation factor in the full model in the overall population. 
Also check liniearity...

## Outcomes

QUESTION: Outcome: cardiovascular death/myocardial infarction/stroke. In stroke is TIA included (so can use the outcome already defined in shfdb3)

The following outcomes are considered: 

- Cardiovascular mortality/first HF hospitalization
- All-cause mortality
- Cardiovascular mortality
- First HF hospitalization
- First cardiovascular hospitalization
- Cardiovascular death/first hospitalization for myocardial infarction or stroke 

The main analysis is the ps matched analysis including NT-proBNP with sglt2 
according to the ITT principle. All analyses are on this population unless 
stated otherwise. The primary endpoint is considered to be 
first HF hospitalization/CVD. 

Incidence per 1000 py was calculated with 95% Poisson confidence intervals. 

The first HF hospitalization/CVD is presented with the 1 - Kaplan-Meier curve
and tested with the log-rank test for the matched population 
(note that the log-rank test is theoretically 
redundant since the cox regression gives the same information). 

QUESTION: KM curves for other outcomes?

QUESTION: 1-KM curves (and corresponding numbers for EVENT (not surv) at 30, 60 days ect) or suvival?

Cox proportional hazards regressions were used to evaluate the association 
between sglt2 and the respective outcomes. Data were censored at 2018-12-31 
or death, so this assumes that the risk of cv mortality and hf hospitalization 
would have been the same for a censored patient had the censoring not occurred. 
In the matched cohort the matched pairs were modeled using a frailty term. 

### Consistency analysis

#### Crossover

In the main analysis sglt2 use is considered according to the ITT principle 
(sglt2 at baseline). As a consistency analysis patients were censored at crossover. 
In untreated patients, crossover was defined as a new sglt2 dispension 
in the Dispensed Drug Registry and for treated patients crossover was defined 
as no sglt2 dispension within 6 months from baseline or from last dispension, 
where crossover was set 3 months from index or last refill. 

QUESTION: We could do a time dependent analysis instead of censoring 
but since problematic to define exact cross-over for users might not be very 
accurate... so perhaps not worth the hassle...

#### Excluding NT-proBNP QUESTION: DO THIS?

Since the % missing observations with NT-proBNP was high, analyses were also 
performed excluding NT-proBNP in the imputation, matching and outcomes analyses 
resulting in N = `r matchp_sens %>% filter(ddr_sglt2 == "Yes") %>% count()`
matched patients with sglt2.  

#### Competing risk

The analyses were also performed using a sub-distributional hazards model [@fg] 
were death was treated as a competing event. In addition, cumulative incidence 
curves are presented were death is treated as a competing event. 

```{r, child = "../src/km.Rmd"}

```

Note to self: N at risk under curve?

```{r, child = "../src/cuminc.Rmd"}

```

```{r, child = "../src/outtab.Rmd"}

```

### Subgroups

QUESTION: Worried about small numbers... I think we should skip subgroup analyses if not VERY important. 

### SGLT2 vs other antidiabetic treatments

QUESTION: Need to discuss based on tab 1. 

### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) for the primary and secondary outcome 
in the ps matched cohort (bbl according to ITT). 
Possible outliers were visually inspected by plotting the dfbetas. 

# Reproducibility

## R session information {#sessioninfo}

`r sessionInfo()$R.version$version.string` [@r] is used, for more details see below.

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: XXX. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
